The human in the loop: volunteer computing as a socio-technical system

### 2016-02-29 Alife16/paper.tex
Hope this is better refs 939f886
First full draft

As far as we know, this paper presents one of the few experiments
that measure the performance of a sociotechnical
metacomputer, that is, a spontaneously created parallel computer that
uses social networks for operations such as gathering new
users. Apol{\'o}nia et al. \citep{apolonia2012enhancing} used the 
Facebook protocol to distribute tasks among the {\em walls} of
friends, explicitly using the social network for computing. However,
it stopped short of relating performance to the macro measures of the
users' social networks. As in the previous example, a social network
was used to get new network nodes; in the previous case a web page was
used, while Facebook's wall was used here.
The algorithms used, as well as the methodology 
for gathering resources will be described next, 
together with the results obtained in this initial setup.

@@

In general, linking and finding correlations between user choices and
performance is an interesting avenue to explore in the future. Even if
the three previous experiments were published in a similar way, one
obtained up to 5 times more total cycles  than the one with the least
number of cycles. It is also essential to obtain volunteers as
simultaneously as possible, so it is possible that the features of the
social network in terms of real-time use will also play a big
role. Even as it is difficult to create controlled experiments in this
area, it is an interesting challenge to explore in the future.

Although we think that, in general, the results presented in this
paper are independent of the problem chosen, it is true that even if
it is a difficult problem for evolutionary algorithms, it takes a few
seconds with the right settings to solve. 
Besides, the two versions of the framework include as an
algorithmic variant using random population size. We do not think this
has had a big influence on the results and in fact this is not noticed
for F15, which is the second version tested but it would be interesting to
measure exactly what this influence has been. In general there are
many issues with the evolutionary algorithm implementation itself,
including using different, or adaptive, policies for inserting and
sending individuals to the pool,
using different policies for population initialization, and also the
incorporation of high-speed local resources to the pool to check what
would be the real influence of the volunteer pool to the final
performance.

Another area is how to enhance the number and quality of
volunteers. For instance, adding a bit of more
control to the user might contribute also to gamification. Right now
the user has only two Web Workers. It is a matter of a single click to
open more tabs in the browser, giving more Web Workers, but it would be
interesting to put everything in a single page and under our control,
to check how often this happens and under which conditions.

Finally, the implementation needs some refinement in terms of
programming and also ease of use. Tools such as Yeoman might be used % ¿añadir una cita o la URL http://yeoman.io/ a pie de página ?
to create a generator in which the user just has to create a fitness
function, with the rest of the framework wrapped around
automatically. All this, as the data used for this paper and the paper
itself, has been published with a free license in GitHub at
\url{https://github.com/JJ/modeling-volunteer-computing}.

>>>>

As far as we know, this paper presents one of the few experiments
that measure the performance of a sociotechnical
metacomputer, that is, a spontaneously created parallel computer that
uses social networks for operations such as gathering new
users. Apol{\'o}nia et al. \citep{apolonia2012enhancing} used the 
Facebook protocol to distribute tasks among the {\em walls} of
friends, explicitly using the social network for computing. However,
it stopped short of relating performance to the macro measures of the
users' social networks. As in the previous example, a social network
was used to get new network nodes; in the previous case a web page was
used, while Facebook's wall was used here.

In our case, social networks are an integral part of the system and
used to spontaneously obtain users. The algorithms used, as well as the methodology 
for gathering resources will be described next, 
together with the results obtained in this initial setup.

@@

In general, linking and finding correlations between user choices and
performance is an interesting avenue to explore in the future. Even if
these  experiments were published in a similar way, one
obtained up to 5 times more total cycles than the one with the least
number of cycles. It is also essential to obtain volunteers as fast and
simultaneously as possible, so it is possible that the features of the
social network in terms of real-time use will also play a big
role; synchronous webs such as Snapchat, which mystifies the writers
of this paper, and Twitter, thanks to its real time nature, might be
better suited than Facebook, LinkedIn or Google plus. Even as it is
difficult to create controlled experiments in this 
area, it is an interesting challenge to explore in the future.

The other area to explore is the algorithmic area itself. Are there
ways to change the evolutionary algorithm, or its visualization, so
that the user has a bigger impact on the result? One of the users in
Twitter even suggested to embed videos so that people spent time
looking at them, but other possible way was to make the user engage
the algorithm by giving him or her buttons to change the mutation rate
when the algorithm is stalled, for instance. If this is combined with
a score board where local performance is compared to other users,
engagement might be increased and thus the performance of the
system. In general there are 
many issues with the evolutionary algorithm implementation itself,
including using different, or adaptive, policies for inserting and
sending individuals to the pool,
using different policies for population initialization, and also the
incorporation of high-speed local resources to the pool to check what
would be the real influence of the volunteer pool to the final
performance.

Finally, the implementation needs some refinement in terms of
programming and also ease of use. Tools such as Yeoman for generating
easily new experiments might be used, so that the user would have  to
create only a fitness
function, with the rest of the framework wrapped around
automatically.

All these avenues of experimentation will be done openly following the
Open Science policy of our group, which, in fact, contributes to
establish trust and security between us and volunteers and is an
essential feature of the system. That is why this paper, as well as
the data and processing scripts, are published with a free license in GitHub at
\url{https://github.com/JJ/modeling-volunteer-computing}.
